---
title: "Northcote Election Data Visualsation notebook"
output: html_notebook
---

# Introduction
This workbook is an album of possible data visualisations for the Northcote Election EDA.

```{r setup}
library(tidyverse)
library(rvest)
library(leaflet)
library(readxl)

library(plotly)

library(rgdal)    # Creates polygon for Northcote District map
```


## Possible List  
- Bar chart of voters by booth by election.  
  - If only 1 year is selected, colour grade the booths by 2pp
- Chart showing the strongest 3 (or so) booths for each candidate relative to other candidates.
- Two-party-preferred timeline.  
- How preferences were allocated.  
  - By party, excluding INDEP
  - By candidate (shows party and years of being candidate after name)
- Map of polling booths, bubble colour showing 2pp
- Timeline of total votes.
  - Also by polling booth
  - By party


# Bar chart of voters by booth by election

```{r bar_chart_voters_by_booth, fig.height=10, fig.width=8}
# elections <- c("2018", "2017", "1999")

bc_booths <-
  bar_chart_voters_by_booth(elections = c("2018", "2017", "1999"))

ggplotly(bc_booths)

```

# Two-party-preferred timeline

```{r two_pp_timeline}

two_pp_detail <- 
  cre_two_pp_detail() %>% 
  glimpse()

```

```{r}
two_pp_all_booth <-
  cre_two_pp_all_booth()

  
plot_two_pp_all_booth()

```

```{r}
two_pp %>% 
  group_by(year, booth) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  spread(key = year, value = votes)
```

```{r}


two_pp_by_booth_nondom <-
  cre_two_pp_by_booth_nondom() %>% 
  glimpse()
```

```{r}
plot_two_pp_by_booth_nondom(p_booth = "Northcote")
```

# Timeline of total votes

```{r votes_by_booth_elec}
votes_by_booth_elec <- 
  cre_votes_by_booth_elec()  

votes_by_booth_all <-
  cre_votes_by_booth_all(votes_by_booth_elec) %>% 
  glimpse()
```
```{r votes_by_elect_ggp}

plot_votes_by_booth(votes_by_booth_all, p_lump_n = 8)

```
```{r votes_by_elect_booth_ggp}

plot_votes_by_booth_all(votes_by_booth_all) 

```

## Timeline of Party votes
```{r party_votes_by_elec}
party_votes_by_elec <-
  cre_party_votes_by_elec(first_pref, elec_dates) %>% 
  glimpse
```
```{r party_votes_by_elec_ggp}

plot_party_votes_by_elec(party_votes_by_elec)

```


# How preferences were allocated.
## By party, excluding INDEP

```{r pref_w_party}
pref_w_party <- 
  cre_pref_w_party(pref, cands_std) %>% 
  glimpse()
  
```

```{r pref_w_party_ggp}

plot_distn_party_prefs("Liberal Democrats", pref_w_party, elec_dates) 

```



## By candidate (shows party and years of being candidate after name)

```{r}

cand_list <-
  cre_cand_list(cands_std) %>% 
  glimpse()


```

```{r pref_w_cand_ggp}

plot_pref_distn_sel_cand("BISHOP, Robert", pref_w_party) 

```

# Map of Polling Booths

```{r votes_by_phys_booth}
votes_by_phys_booth <-
  cre_votes_by_phys_booth(votes_by_booth_all, booth_addr_lst) %>% 
  # print()
  glimpse

```


```{r}

print_booth_map(votes_by_phys_booth, "2017")
```

