---
title: "Data Visualisation for EDA"
output: html_notebook
---

Assumes that the data is already ingested, according to ingest_data.Rmd.  



```{r Vote source}
first_pref_tbl %>% 
  group_by(poll_src) %>% 
  summarise(votes_sum = sum(votes, na.rm = TRUE)) %>% 
  mutate(poll_src = poll_src %>% fct_reorder(votes_sum)) %>% 
  
  ggplot(aes(x = poll_src, y = votes_sum)) +
  geom_col() +
  scale_y_continuous(breaks = seq(0, 10000, 1000)) +
  coord_flip()

```

Early votes and postal votes are by far the most frequent source.  

Northcote South is remarkably common given that it is on the fringe of the district.  This may be due to the proximity of public and higher density housing.  

```{r}
first_pref_tbl %>% 
  group_by(cand) %>% 
  mutate(poll_src_sh = votes / sum(votes, na.rm = TRUE),
         cand_votes = str_c(cand, "; ", sum(votes, na.rm = TRUE), sep = ""),
         vote_sum = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(cand_votes = cand_votes %>% fct_reorder(vote_sum),
         poll_src = poll_src %>% fct_reorder(votes)) %>% 
  
  ggplot(aes(x = cand_votes, y = poll_src_sh, fill = poll_src)) +
  geom_col(position = "fill", colour = "grey") +
  
  coord_flip()

```
Clare Burns' heartland appears as Northcote North, whereas Lidia Thorp's votes came from a much wider source.

Vince Fontana was very strong in Early Votes, which possibly reflects his high profile early in the campaign.  

Would be good to use plotly with this.

```{r}

# library(plotly)

first_pref_tbl %>% 
  group_by(cand) %>% 
  mutate(poll_src_sh = votes / sum(votes, na.rm = TRUE),
         cand_votes = str_c(cand, "; ", sum(votes, na.rm = TRUE), sep = ""),
         vote_sum = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(cand_votes = cand_votes %>% fct_reorder(vote_sum),
         poll_src = poll_src %>% fct_reorder(votes),
         hover_txt = str_c(cand, ", ", poll_src, ": ", scales::percent(poll_src_sh), sep = "")) %>% 
  # print()
  
  plot_ly(y = ~cand_votes, x = ~poll_src_sh, text = ~hover_txt, 
          hoverinfo = "text",
          type = 'bar', name = ~poll_src, orientation = 'h') %>%
  layout(xaxis = list(title = 'Votes source (cum)'), yaxis = list(title = ""), barmode = 'stack')

```

```{r}

dist_prog_votes %>% 
  ggplot(aes(x = rnd_num, y = cand)) +
  geom_point(aes(size = votes, colour = votes_ch), na.rm = TRUE) +
  # scale_size(trans = "log10", breaks = c(10, 100, 500, 1000, 3000, 5000)) +
  scale_colour_gradient(trans = "log10")

```

