---
title: "Northcote Election Data Visualsation notebook"
output: html_notebook
---

# Introduction
This workbook is an album of possible data visualisations for the Northcote Election EDA.

```{r setup}
library(tidyverse)
library(rvest)
library(leaflet)
library(readxl)

library(plotly)

library(rgdal)    # Creates polygon for Northcote District map
```


## Possible List  
- Bar chart of voters by booth by election.  
  - If only 1 year is selected, colour grade the booths by 2pp
- Chart showing the strongest 3 (or so) booths for each candidate relative to other candidates.
- Two-party-preferred timeline.  
- How preferences were allocated.  
  - By party, excluding INDEP
  - By candidate (shows party and years of being candidate after name)
- Map of polling booths, bubble colour showing 2pp
- Timeline of total votes.
  - Also by polling booth
  - By party


# Bar chart of voters by booth by election

```{r bar_chart_voters_by_booth, fig.height=10, fig.width=8}
# elections <- c("2018", "2017", "1999")

bc_booths <-
  bar_chart_voters_by_booth(elections = c("2018", "2017", "1999"))

ggplotly(bc_booths)

```

# Two-party-preferred timeline

```{r two_pp_timeline}
booths_exclude <- c("Ordinary Votes Total")

two_pp_detail <- 
  two_pp %>% 
  filter(!(booth %in% booths_exclude)) %>% 
  # join to party
  left_join(cands_std %>% select(candidate, party_std, year), by = c("candidate", "year")) %>% 
  # join to date
  left_join(elec_dates %>% select(elec_ID, date), by = c("year" = "elec_ID")) %>% 
  glimpse()

```

```{r}
two_pp_all_booth <-
  two_pp_detail %>% 
  mutate(party_std = coalesce(party_std, "Informal")) %>% 
  group_by(party_std, date) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # complete(party_std, date) %>% 
  group_by(date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(non_dom_party = case_when(
    is.null(party_std) ~ FALSE,
    party_std == dom_party_name ~ FALSE,
    party_std == "Informal" ~ FALSE,
    TRUE ~ TRUE
  ),
  booth = "All booths",
  hov_text = str_c(booth, "<br>", party_std, "<br>2pp: ", scales::percent(votes_sh)))

  
two_pp_plot <- 
  two_pp_all_booth %>% 
  ggplot(aes(x = date, y = votes_sh, colour = party_std, shape = party_std)) +
  geom_line(na.rm = TRUE) +
  geom_point(na.rm = TRUE, aes(text = hov_text)) +
  geom_hline(yintercept = 0.5, linetype = 2, size = 0.5, colour = "grey") +
  scale_y_continuous("2 party preferred share", labels = scales::percent) +
  scale_color_manual("Party", values = party_colours()) +
  scale_shape_discrete("Party") +
  labs(title = "Two party preferred share", x = "")

ggplotly(two_pp_plot, tooltip = "text")

```

```{r}
two_pp %>% 
  group_by(year, booth) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  spread(key = year, value = votes)
```

```{r}
dom_party_name <- "Australian Labor Party"

two_pp_by_booth_nondom <-
  # tibble with the non-dominant party
  two_pp_detail %>% 
  group_by(booth, date, party_std) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  group_by(booth, date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!is.na(party_std)) %>% 
    # remove informal votes from chart
  filter(party_std != dom_party_name) %>% 
  # semi_join(booth_addr_lst, by = "booth") %>% 
  mutate(hov_text = str_c(booth, "<br>", party_std, "<br>", scales::percent(votes_sh))) %>%
  glimpse()
```

```{r}

two_pp_by_booth_ggplot <-
  two_pp_by_booth_nondom %>% 
  semi_join(booth_addr_lst, by = "booth") %>% 
  # mutate(booth = factor(booth)) %>% 
  ggplot(aes(x = date, y = votes_sh, shape = party_std, group = booth)) +
  geom_line(na.rm = TRUE, colour = "grey", aes(group = booth, shape = party_std)) +
  geom_point(na.rm = TRUE, colour = "grey", size = 0.5, aes(text = hov_text)) +
  geom_hline(yintercept = 0.5, linetype = 2, size = 0.5, colour = "grey") +
  scale_y_continuous("2 party preferred share", labels = scales::percent) +
  # scale_shape_manual("Booth", values = 97:111) +
  # scale_shape_identity() +
  geom_line(data = two_pp_by_booth_nondom %>% filter(booth == "Westgarth"), colour = "black") +
  geom_point(data = two_pp_by_booth_nondom %>% filter(booth == "Westgarth"), aes(text = hov_text, colour = party_std)) +
  scale_color_manual("Party", values = party_colours()) +
  scale_shape_discrete("Party") +
  geom_line(data = two_pp_all_booth %>% filter(non_dom_party), colour = "blue") +
  geom_point(data = two_pp_all_booth %>% filter(non_dom_party), aes(text = hov_text, colour = party_std)) +
  labs(title = "Two party preferred share by booth", x = "",
       caption = "Westgarth highlighted. Hover mouse for more detail")


ggplotly(two_pp_by_booth_ggplot, tooltip = "hov_text")
```

# Timeline of total votes

```{r votes_by_booth_elec}
votes_by_booth_elec <- 
  first_pref %>% 
  inner_join(elec_dates %>% select(year = elec_ID, date), by = "year") %>% 
  group_by(date, booth, year) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  left_join(two_pp_by_booth_nondom %>% select(-votes), by = c("date", "booth")) %>% 
  filter(!is.na(party_std)) 

votes_by_booth_all <-
  votes_by_booth_elec %>% 
  filter(votes != 0) %>% 
  group_by(date, year, party_std) %>% 
  summarise(booth = "All",
            votes_sh = sum(votes_sh * votes / sum(votes, na.rm = TRUE)),
            votes = sum(votes, na.rm = TRUE),
            hov_text = NA_character_) %>% 
  ungroup() %>% 
  select(names(votes_by_booth_elec)) %>% 
  union(votes_by_booth_elec) %>% 
  filter(votes != 0) %>% 
  # mutate(booth_ggp = fct_lump(booth, 8, w = votes)) %>% 
  glimpse()
```
```{r votes_by_elect_ggp}
votes_by_elect_ggp <-
  votes_by_booth_all %>% 
  mutate(booth_ggp = fct_lump(booth, 8, w = votes) %>% fct_relevel("All")) %>% 
  group_by(date, booth_ggp, year) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  group_by(date) %>% 
  mutate(votes_sh = votes / sum(if_else(booth_ggp == "All", votes, 0), na.rm = TRUE)) %>% 
  ungroup %>% 
  mutate(hov_text = str_c(booth_ggp, ": ", scales::comma(votes), "<br>Share of All: ", scales::percent(votes_sh))) %>% 

  ggplot(aes(x = date, y = votes)) +
  # scale_y_continuous("Votes", labels = scales::comma) +
  scale_y_log10("Votes (log scale)", labels = scales::comma) +
  scale_shape_discrete("Polling Booth") +
  expand_limits(y = 0) +
  geom_point(aes(text = hov_text, shape = booth_ggp)) +
  geom_line(colour = "grey", size = 0.25, aes(group = booth_ggp)) +
  geom_line(data = votes_by_booth_all %>% filter(booth == "All"), colour = "red", size = 0.5) +
  labs(title = "Total votes", x = "")
  
ggplotly(votes_by_elect_ggp, tooltip = "text")
  
```
```{r votes_by_elect_booth_ggp}
votes_by_elect_booth_ggp <-
  votes_by_booth_all %>% 
  filter(booth != "All") %>% 
  group_by(date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE)) %>% 
  ungroup %>% 
  mutate(hov_text = str_c(booth, ": ", scales::comma(votes), "<br>Share of total: ", scales::percent(votes_sh))) %>% 
  
  ggplot(aes(x = date, y = votes)) +
  scale_y_log10("Votes (log scale)", labels = scales::comma) +
  expand_limits(y = 0) +
  geom_point(aes(text = hov_text), size = 0.25) +
  geom_line(colour = "grey", size = 0.25, aes(group = booth)) +
  geom_line(data = votes_by_booth_all %>% filter(booth == "Westgarth"), colour = "red", size = 0.5) +
  labs(title = "Total votes by booth", x = "")
  
ggplotly(votes_by_elect_booth_ggp, tooltip = "text")
```

## Timeline of Party votes
```{r party_votes_by_elec}
party_votes_by_elec <-
  first_pref %>% 
  inner_join(elec_dates %>% select(year = elec_ID, date), by = "year") %>% 
  group_by(date, cand_party) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(cand_party = cand_party %>% str_replace_all(., "\\'", "") %>% str_trim()) %>% 
  left_join(party_std, by = "cand_party") %>% 
  glimpse
```
```{r party_votes_by_elec_ggp}
party_votes_by_elec_ggp <-
  party_votes_by_elec %>% 
  group_by(date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE),
         party_std = coalesce(party_std, "Informal") %>% fct_relevel("Informal", after = Inf),
         hov_text = str_c(party_std, "<br>Votes: ", scales::comma(votes), " (", scales::percent(votes_sh), ")")) %>% 
  ungroup %>% 
  # mutate(party_std = coalesce(party_std, "Informal") %>% fct_relevel("Informal", after = Inf)) %>% 
  # glimpse

  ggplot(aes(x = date, y = votes)) +
  geom_point(aes(text = hov_text, colour = party_std, fill = party_std), size = 0.5) +
  geom_line(aes(group = party_std, colour = party_std), size = 0.25) +
    scale_fill_manual("Party", values = party_colours()) +
  scale_colour_manual("Party", values = party_colours()) +
  scale_y_log10(label = scales::comma) +
  labs(title = "Total votes by party", x = "") +
  expand_limits(y = 0) +
  geom_line(data = votes_by_booth_all %>% filter(booth == "All"), colour = "grey", size = 0.25, aes(group = booth, text = "Total"))

ggplotly(party_votes_by_elec_ggp, tooltip = "text")
  
```


# How preferences were allocated.
## By party, excluding INDEP

```{r pref_w_party}
pref_w_party <- 
  pref %>% 
  left_join(cands_std %>% select(-cand_party) %>% rename_all(~str_c(., ".from")), 
            by = c("from_cand" = "candidate.from", "year" = "year.from")) %>% 
  left_join(cands_std %>% select(-cand_party) %>% rename_all(~str_c(., ".to")), 
            by = c("to_cand" = "candidate.to", "year" = "year.to")) %>% 
  glimpse()
  
```

```{r pref_w_party_ggp}
pref_w_party_ggp <-
  pref_w_party %>% 
  complete(year, party_std.from) %>% 
  filter(party_std.from == "Fiona Pattens Reason Party (Sex Party)") %>% 
  mutate(hovtext = str_c(party_std.to, " (", to_cand, ")<br>", votes, " votes. (", scales::percent(votes / votes_sum), ")")) %>% 
  # glimpse()

  ggplot(aes(x = year, y = votes)) +
  geom_bar(aes(fill = party_std.to, text = hovtext), stat = "identity", position = "dodge") +
  scale_fill_manual("Party", values = party_colours()) +
  
  labs(title = "Distribution of Liberal votes by election", x = "", y = "votes")

ggplotly(pref_w_party_ggp, tooltip = "text")

```



## By candidate (shows party and years of being candidate after name)

```{r}
cands_std %>% 
  filter(year >= "2010") %>% 
  group_by(candidate) %>% 
  summarise(year_num = n(),
            years = str_c(year, collapse = ","),
            parties = str_c(unique(party_std), collapse = ",")) %>% 
  ungroup %>% 
  arrange(-year_num)
```

```{r pref_w_cand_ggp}
pref_w_cand_ggp <- 
  pref_w_party %>% 
  filter(from_cand == "EDWARDS, Bryony") %>% 
  mutate(hovtext = str_c(party_std.to, " (", to_cand, ")<br>", votes, " votes. (", scales::percent(votes / votes_sum), ")")) %>% 
  # glimpse()
  
  ggplot(aes(x = year, y = votes)) +
  geom_bar(aes(fill = party_std.to, text = hovtext), stat = "identity", position = "dodge") +
  scale_fill_manual("Party", values = party_colours()) +
  
  labs(title = "Distribution of Liberal votes by election", x = "", y = "votes")  

ggplotly(pref_w_cand_ggp, tooltip = "text")
```

