---
title: "Northcote Election Data Visualsation notebook"
output: html_notebook
---

# Introduction
This workbook is an album of possible data visualisations for the Northcote Election EDA.

```{r setup}
library(tidyverse)
library(rvest)
library(leaflet)
library(readxl)

library(plotly)

library(rgdal)    # Creates polygon for Northcote District map
```


## Possible List  
- Bar chart of voters by booth by election.  
- Chart showing the strongest 3 (or so) booths for each candidate relative to other candidates.
- Two-party-preferred timeline.  
- How preferences were allocated


# Bar chart of voters by booth by election

```{r bar_chart_voters_by_booth, fig.height=10, fig.width=8}
# elections <- c("2018", "2017", "1999")

bc_booths <-
  bar_chart_voters_by_booth(elections = c("2018", "2017", "1999"))

ggplotly(bc_booths)

```

# Two-party-preferred timeline

```{r two_pp_timeline}
booths_exclude <- c("Ordinary Votes Total")

two_pp_detail <- 
  two_pp %>% 
  filter(!(booth %in% booths_exclude)) %>% 
  # join to party
  left_join(cands_std %>% select(candidate, party_std, year), by = c("candidate", "year")) %>% 
  # join to date
  left_join(elec_dates %>% select(elec_ID, date), by = c("year" = "elec_ID")) %>% 
  glimpse()

```

```{r}
two_pp_all_booth <-
  two_pp_detail %>% 
  mutate(party_std = coalesce(party_std, "Informal")) %>% 
  group_by(party_std, date) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # complete(party_std, date) %>% 
  group_by(date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(non_dom_party = case_when(
    is.null(party_std) ~ FALSE,
    party_std == dom_party_name ~ FALSE,
    party_std == "Informal" ~ FALSE,
    TRUE ~ TRUE
  ))

  
two_pp_plot <- 
  two_pp_all_booth %>% 
  ggplot(aes(x = date, y = votes_sh, colour = party_std, shape = party_std)) +
  geom_line(na.rm = TRUE) +
  geom_point(na.rm = TRUE) +
  geom_hline(yintercept = 0.5, linetype = 2, size = 0.5, colour = "grey") +
  scale_y_continuous("2 party preferred share", labels = scales::percent) +
  scale_color_manual("Party", values = party_colours()) +
  scale_shape_discrete("Party") +
  labs(title = "Two party preferred share", x = "")

ggplotly(two_pp_plot)

```

```{r}
two_pp %>% 
  group_by(year, booth) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  spread(key = year, value = votes)
```

```{r}
dom_party_name <- "Australian Labor Party"

two_pp_by_booth_nondom <-
  # tibble with the non-dominant party
  two_pp_detail %>% 
  group_by(booth, date, party_std) %>% 
  summarise(votes = sum(votes, na.rm = TRUE)) %>% 
  group_by(booth, date) %>% 
  mutate(votes_sh = votes / sum(votes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!is.na(party_std)) %>% 
    # remove informal votes from chart
  filter(party_std != dom_party_name) %>% 
  semi_join(booth_addr_lst, by = "booth") %>% 
  glimpse()
```

```{r}

two_pp_by_booth_ggplot <-
  two_pp_by_booth_nondom %>% 
  # mutate(booth = factor(booth)) %>% 
  ggplot(aes(x = date, y = votes_sh)) +
  geom_line(aes(shape = booth), na.rm = TRUE, colour = "grey") +
  geom_point(aes(shape = booth), na.rm = TRUE, colour = "grey") +
  geom_hline(yintercept = 0.5, linetype = 2, size = 0.5, colour = "grey") +
  scale_y_continuous("2 party preferred share", labels = scales::percent) +
  scale_shape_manual("Booth", values = 97:111) +
  # scale_shape_identity() +
  labs(title = "Two party preferred share", x = "") +
  geom_line(data = two_pp_by_booth_nondom %>% filter(booth == "Westgarth"), colour = "black") +
  geom_point(data = two_pp_by_booth_nondom %>% filter(booth == "Westgarth"), colour = "black") +
  geom_line(data = two_pp_all_booth %>% filter(non_dom_party), colour = "blue")


ggplotly(two_pp_by_booth_ggplot)
```

